---
title: "HW6"
author: "Jiayi Shen"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      dpi = 300)
library(tidyverse)
library(rvest)

```


#Problem 1 
```{r load homicide data, message = FALSE}
homicide_data = read_csv("./data/homicide-data.csv")
```

```{r homicide data tidying part 1}
#Create a city_state variable
homicide_data =
homicide_data %>% 
  mutate(city_state = str_c(city, ",", state)) %>% 
  select(-city, -state)

# create a binary variable indicating whether the homicide is solved. 
homicide_data =
homicide_data %>% 
  filter(!city_state %in% c("Dallas,TX", "Phoenix,AZ", "Kansas City,MO", "Tulsa,AL")) %>% 
  mutate(solved =as.numeric(disposition == "Closed by arrest"))
  
```


```{r homicide data tidying part 2}
# modifiy victim_race to have categories white and non-white 
# with white as the reference category
homicide_data = 
  homicide_data %>% 
  mutate(victim_race = as_factor(ifelse(victim_race == "White", "white","non-white"))) %>% 
  mutate(victim_race = fct_relevel(victim_race, "white")) %>% 
  mutate(victim_age = as.numeric(victim_age)) 
```


```{r logistic regression to homicide data, message = FALSE}
# fit a logistic regression with "resolved" as the outcome and victim age, sex and race as predictors
fit_logistic = 
  homicide_data %>% 
  filter(city_state == "Baltimore,MD") %>% 
  glm(solved ~ victim_age + victim_sex + victim_race, data = ., family = binomial()) 


# obtain the estimate and confidence interval of the adjusted odds ratio
tibble1 = 
fit_logistic %>% 
  broom::tidy() %>% 
  mutate(OR_estimate = exp(estimate)) %>% 
  select(term, OR_estimate, p.value)

tibble2 = as_tibble(exp(confint.default(fit_logistic)))

cbind(tibble1, tibble2) %>% knitr::kable(digits = 3)
```

**Compare non-white victims to white victims while keeping all other variables fixed:**
According to the result of `glm`, the estimates of adjusted odds ratio for the predictor `victim_race`(non-white compared to white victims) is less than 0.5 with a p-value less than 0.01. Thus we conclude that in Baltimore, homicide cases with non-white victims are significantly less likely to be resolved than those with white victims, given that all other variables are fixed. 


```{r glm for each of the cities}
glm_result = function(df){
  fit = 
  df %>% 
  filter(!victim_sex == "Unknown") %>% 
  glm(solved ~ victim_age + victim_sex + victim_race, data = ., family = binomial()) 

  t1 = 
  fit %>% 
    broom::tidy() %>% 
    mutate(OR_estimate = exp(estimate)) %>% 
    select(term, OR_estimate, p.value)

  t2 = as_tibble(exp(confint.default(fit)))
  
  return(cbind(t1, t2))
}


# run glm for each of the cities in your dataset
homicide_data %>% 
  select(city_state, solved, victim_age, victim_sex, victim_race) %>% 
  nest(solved:victim_race) %>%
  mutate(glm = map(data, glm_result)) %>% 
  select(-data) %>% 
  unnest() %>% 
  filter(term == "victim_racenon-white") # obtain coef only for the predictor "race"


```

